# Introduction {#sec-intro}

This chapter demonstrates dynamic R computation, data exploration, and cross-referencing.

## Exploring Motor Trend Data

We analyze the classic `mtcars` dataset. First, we inspect a styled summary table shown in @tbl-cars.

```{r}
#| label: tbl-cars
#| tbl-cap: "Top 5 rows of Motor Trend Car Road Tests"
#| echo: true

knitr::kable(head(mtcars[, 1:4], 5))
```

Next, we inspect the relationship between engine displacement and fuel efficiency in @fig-mtcars-plot.

```{r}
#| label: fig-mtcars-plot
#| fig-cap: "Displacement vs. Miles Per Gallon (MPG)"
#| fig-align: "center"
#| out-width: "85%"
#| echo: true

plot(
  mtcars$disp, mtcars$mpg,
  xlab = "Displacement (cu.in.)",
  ylab = "Miles Per Gallon (MPG)",
  pch = 19,
  col = "steelblue"
)
abline(
  lm(mpg ~ disp, data = mtcars),
  col = "firebrick",
  lwd = 2
)
```

As demonstrated in @fig-mtcars-plot and @sec-intro, vehicles with larger engine displacement yield lower fuel economy.

## Professional Styling Features

### Callout Boxes

Technical books use callouts to highlight important reader notes:

\Needspace{6\baselineskip}
::: {.callout-note}
## Important Note for Readers

The `mtcars` dataset was extracted from the 1974 *Motor Trend* US magazine, comprising fuel consumption and 10 aspects of automobile design.
:::

\Needspace{6\baselineskip}
::: {.callout-warning}
## Common Pitfall

Never evaluate models without checking for missing values (`NA`) or extreme outliers first!
:::

### Hiding Code but Showing Results

Publishers often want clean graphs without printing the underlying script in the main text:

```{r}
#| label: fig-hidden-code
#| fig-cap: "Fuel Efficiency Distribution (Code Hidden)"
#| fig-align: "center"
#| echo: false

hist(
  mtcars$mpg,
  col = "cornflowerblue",
  border = "white",
  main = "",
  xlab = "Miles Per Gallon (MPG)"
)
```
