# Introduction to data {#sec-intro}

Scientists seek to answer questions using rigorous methods and careful observations [@sec-intro]. Data collected from field notes, surveys, and experiments form the backbone of statistical investigations.

## Data basics

Effective organization and description of data is a first step in most analyses.

### Observations, variables, and data matrices

A **data matrix** organizes cases as rows and characteristics as columns.

```{r}
#| label: tbl-loans
#| tbl-cap: "Sample records from a peer-to-peer lending portfolio"
#| echo: false

loan_sample <- data.frame(
  loan_amount = c(22000, 6000, 25000, 15000),
  interest_rate = c(10.90, 9.92, 26.30, 6.08),
  term = c(60, 36, 36, 36),
  grade = c("B", "B", "E", "A"),
  homeownership = c("rent", "rent", "mortgage", "mortgage")
)

knitr::kable(loan_sample, booktabs = TRUE)
```

As illustrated in @tbl-loans, each row corresponds to an observational unit.

::: {.callout-note title="DATA MATRIX"}
Each row of a data matrix corresponds to a unique case (observational unit), and each column corresponds to a variable.
:::

::: {.callout-tip title="EXAMPLE 1.1: CLASSIFYING VARIABLES"}
**Example problem:** In the loan sample shown above, classify `loan_amount` and `homeownership` as either numerical or categorical.

**Solution to the example:** `loan_amount` takes numerical dollar values where arithmetic is sensible, so it is numerical. `homeownership` categorizes borrowers into distinct groups (rent or mortgage), making it categorical.
:::

::: {.callout-warning title="GUIDED PRACTICE 1.2: PROPORTION OF RENTERS"}
What proportion of the loans shown in @tbl-loans represent borrowers who rent their homes?^[Two out of four loans in the sample represent renters: $2/4 = 0.50$ or $50\%$.]
:::

## Scatterplots and Associations

Scatterplots provide a case-by-case view of data for two numerical variables.

```{r}
#| label: fig-scatter
#| fig-cap: "Engine displacement versus fuel economy for passenger vehicles"
#| fig-width: 6.5
#| fig-height: 3.5
#| echo: false

plot(
  mtcars$disp, mtcars$mpg,
  xlab = "Engine Displacement (cu.in.)",
  ylab = "Miles Per Gallon (MPG)",
  pch = 19, col = "#2E5B88"
)
abline(lm(mpg ~ disp, data = mtcars), col = "#B22222", lwd = 2)
```

As demonstrated in @fig-scatter, there is a clear downward trend: vehicles with larger engine displacement yield lower fuel economy.

## Exercises

1.1 **Sampling methods.** A survey selects 100 students at random from a university directory.
(a) Identify the target population and the sample.
(b) Is this an observational study or an experiment?